---
title: "EDDA Assignment 2"
author: "Agnes Nikki Tim, group 38"
date: "13 February 2023"
output: pdf_document
fontsize: 11pt
highlight: tango
---

# Exercise 1: Trees

### The *Amsterdamsche Bos* forestry wishes to estimate the total wood volume of the trees on its domain. To this end the forestry has cut a sample of 59 trees of their most prevalent types beech and oak and collected the data from the cut trees in the file [treeVolume.txt](https://canvas.vu.nl/courses/68030/files/6051344?wrap=1 "treeVolume.txt")[Download treeVolume.txt](https://canvas.vu.nl/courses/68030/files/6051344/download?download_frd=1). The volume of these trees alongside with their height and trunk diameter have been measured; these are the columns *volume*, *height* and *diameter*, respectively. Column *type* gives the tree type: *Beech* or *Oak*. The tree type, height and diameter can be measured in the field without sacrificing the tree. The forestry hypothesizes that these are predictive of the wood volume.

## A

### Investigate whether the tree type influences volume by performing ANOVA, without taking diameter and height into account. Can a t-test be related to the above ANOVA test? Estimate the volumes for the two tree types.

```{r}
trees = read.table("treeVolume.txt")
trees
```

```{r, echo=F}
birthweight <- read.table("birthweight.txt", header=TRUE)

```
