---
title: "EDDA Assignment 2"
author: "Agnes Nikki Tim, group 38"
date: "13 February 2023"
output: pdf_document
fontsize: 11pt
highlight: tango
---

# Exercise 1: Military Coups

### To study the influence of different political and geographical variables on the number of military coups, these data are collected for several countries in the file [coups.txt](https://canvas.vu.nl/courses/68030/files/6051349?wrap=1 "coups.txt")

### Perform Poisson regression on the full data set, taking *miltcoup* as response variable. Comment on your findings.

```{r}
coups <- read.table("coups.txt", header=TRUE)
coups
```

Check model assumptions:

```{r}
pairs(coups)
```

```{r}
summary(glm(miltcoup ~ ., data = coups, family = "poisson"))
```

To study the influence of different political and geographical variables on the number of military coups, poisson regression was performed.

The significant coefficient variables were: oligarchy, pollib and parties.

### Use the step-down approach (using output of the function *summary*) to reduce the number of explanatory variables. Compare the resulting model with your findings in a).

For the step-down technique, the variable with the highest p-value is removed sequentially. The first variable to be removed is numelec.

```{r}
summary(glm(miltcoup ~ oligarchy + pollib + parties + pctvote + popn + size + numregim, 
            data = coups, family = "poisson"))
```

The next variable to remove is numregim.

```{r}
summary(glm(miltcoup ~ oligarchy + pollib + parties + pctvote + popn + size, 
            data = coups, family = "poisson"))
```

The next variable to remove is size.

```{r}
summary(glm(miltcoup ~ oligarchy + pollib + parties + pctvote + popn, 
            data = coups, family = "poisson"))
```

The next variable to remove is popn.

```{r}
summary(glm(miltcoup ~ oligarchy + pollib + parties + pctvote, 
            data = coups, family = "poisson"))
```

The next variable to remove is pctvote.

```{r}
summary(glm(miltcoup ~ oligarchy + pollib + parties, 
            data = coups, family = "poisson"))
```

All remaining variables are significant.

### Using the model from b), predict the number of coups for a hypothetical country for all the three levels of political liberalization and the (overall) averages of all the other (numerical) characteristics. Comment on your findings.
