---
title: "EDDA Assignment 2"
author: "Agnes Nikki Tim, group 38"
date: "13 February 2023"
output: pdf_document
fontsize: 11pt
highlight: tango
---

# Exercise 3: Titanic

### Study the data and give a few (\>1) summaries (graphics or tables).

```{r}
titanic = read.table("titanic.txt", header=T)
titanic
```

```{r}
ggplot(data = titanic, aes(x = factor(Age), fill = factor(Survived))) + geom_bar(position = "dodge")
```

### ![]()

```{r}
ggplot(titanic, aes(Sex, fill = factor(Survived))) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent_format())
```

```{r}
# Distribution of survivors by passenger class
ggplot(titanic, aes(PClass, fill = factor(Survived))) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent_format())
```

### Fit a logistic regression model (no interactions yet) to investigate the association between the survival status and the predictors *PClass*, *Age* and *Sex*.

```{r}
# Fit logistic regression model
titanic_model <- glm(Survived ~ PClass + Age + Sex, data = titanic, family = binomial)

# Model summary
summary(titanic_model)

```

### Interpret the results in terms of odds, comment.

...

### Investigate the interaction of predictor *Age* with *PClass*, and the interaction of *Age* with *Sex*. From this and a), choose (and justify) a resulting model.

```{r}
# Fit extended logistic regression model
titanic_model2 = glm(Survived ~ PClass * Age + Age * Sex, data = titanic, family = binomial)
summary(titanic_model2)
```

ASK QUESTION ABOUT THIS

We choose the last model.

### For this model, report the estimate for the probability of survival for each combination of levels of the factors *PClass* and *Sex* for a person of age 55.

```{r}
summary(titanic_model2)$coefficients
```

### Propose a method to predict the survival status and a quality measure for your prediction and describe how you would implement that method (you do not need to implement it).

### Another approach would be to apply a contingency table test and to investigate whether factor passenger class has an effect on the survival status and whether factor gender has an effect on the survival status. Implement the relevant test(s).

```{r}
# Create a contingency table of Pclass and Survived
cont1 = table(titanic$PClass, titanic$Survived)
cont
```

```{r}
chisq.test(cont1)
```

```{r}
# Create a contingency table of Sex and Survived
cont2 = table(titanic$Sex, titanic$Survived)
cont2
```

```{r}
chisq.test(cont2)
```

### Is the second approach in d) wrong? Name both an advantage and a disadvantage of the two approaches, relative to each other.

agnes weet antwoord
